Fictitious Play in Network Aggregative Games

### 2021-05-20 Documents/FP in NAGs/neurips_2020.tex
minor mods

@@

NA games are a recent \cite{Parise2015} extension of
%introduction  which presents
%a variant of
aggregative games by adding an underlying structure to
the agent population. Since its introduction, distributed algorithms have
been built with the aim of finding NE in NA games. In particular,
\cite{Parise2015,Parise2020} consider the case in which payoffs are
given by Lipschitz functions with unique minimisers and apply standard
topological fixed-point towards designing algorithms that
converge to NE. We note that the \emph{Mann Iteration}, one of the
algorithms applied in \cite{Parise2020} shares a remarkably similar
form with the discrete variant of FP. To the best of our knowledge,
this link has not been explored in the literature and is an avenue for
research which we are currently pursuing. Another common algorithm for
distributed NE seeking is the projected gradient (resp. subgradient)
dynamics which is explored in \cite{Zhang2020}
(resp. \cite{Shokri2020, Shokri2021}). In all of these works, the cost
function is assumed to be convex, and therefore have a unique
minimiser. In fact this is a common assumption in works concerning NA
Games \cite{Zhu2021, Lei2020}, which we believe is due to its ubiquity
in control settings. We have not yet come across works which consider
NA games from the point of view of payoff matrices, which are more
common in multi agent learning settings.  Furthermore, to the best of
our knowledge, this is the first work which introduces the application
of a strict learning algorithm in the NA setting.

\paragraph{Fictitious Play}

Fictitious Play was introduced by Brown in 1951 \cite{BrownPublished, BrownUnpublished} as a
'natural' way in which to approximate the Nash Equilibrium in zero sum games. It was known at the
time that a discrete variant of the algorithm would converge to the NE in two player zero-sum games
\cite{Robinson}. Since then a number of proofs of convergence were determined in 2 player games
\cite{Miyasawa, Polak, Berger, Monderer and Sela, Monderer and Shapley}, as well as corresponding
convergence rates \cite{Harris, Shapiro}. Many results hold for both the discrete and continuous
variants of FP. Their relation was finally solved by Hofbauer \cite{Hofbauer}, who showed that
attractors of the discrete time process are also attractors for the continuous time variant. As
such, most of the analysis on FP focuses on the continuous time variant \cite{Ostrovski}. It was
later shown in \cite{Shapley} that there is a family of games (now called the \emph{Shapley family})
for which FP does not converge at all but rather shows periodic behaviour. This led to research
which looks at this non convergent behaviour, specifically showing the existence of more periodic
orbits as well as chaotic behaviour \cite{Sparrow}. However, the work looking at fictitious play
where there are more than two agents is sparse. \cite{Sela} considers a game in which a multi-player
game is decomposed by requiring that each agent engage in a two person game against each of the
opponent. Their payoff is given by the sum of payoffs in all of these subgames. It was found that,
if this game is zero sum, then FP will converge. Similar results for more than two players were
found for games with identical interests in \cite{MondererShapley1994}. In \cite{FPMFG}, the action
of FP was considered in a Mean Field Game, and it was found again that the algorithm converges for
zero sum games. The most general result, and the one most similar to our own, was found by Ewerhart
and Valkanova in 2020 \cite{Ewerhart} in which they showed that fictitious play converges in network
games, where each agent is engaged in a two-player game with each of their neighbours. The rate of
convergence was found to be independent of the network size.  Our work extends the analysis of FP in
multiplayer games by considering its action in NA Games, so that the agents do not play individual
games against each of their neighbours but rather a single game against the aggregate of their
neighbours. In particular we also extend a result for two player games found by Ostrovski and van
Strein \cite{Payoff Performance} which showed that FP achieves no-regret to the multi-player
setting.

\section{Preliminaries}
\subsection{Network Aggregative Games}
\label{sec::NAG}

  The model we consider is that there is a set $\agentset = {1,
           \ldots , N}$ of agents who are connected through an underlying
          interaction graph. This graph is given by the tuple
          $(\agentset, (\edgeset, \weightset))$ in which $\edgeset$ is
          the set of all pairs $(\mu, \nu) \in \agentset \times \agentset$
          connected. Formally,

  \begin{definition}[Interaction Graph]
    Given a set $\agentset$ of agents, an {\em interaction graph} $I = (\agentset, (\edgeset,
    \weightset))$ such that
    \begin{itemize}
    \item $\edgeset \subseteq \agentset \times \agentset$.        Then, the set of neighbours of agent
        $\mu$ is denoted as $N^\mu = \{\nu \in \agentset \, : \, (\mu, \nu) \in \edgeset\}$.
      \item $\weightset \in M_N(\mathbb{R})$ is the weighted adjacency matrix whose elements $w^{\mu
        \nu}$ expresses the importance that agent $\mu$ places on agent $\nu$. If $(\mu, \nu) \not
        \in \edgeset$ then $w^{\mu \nu} = 0$;        $\wmunu \in (0, 1]$ otherwise.
    \end{itemize}
  \end{definition}

  With the above definition, we can formalise the network aggregative (NA) game:

  \begin{definition}[Network Aggregative Game]
    A Network Aggregative game is defined as the tuple $\Gamma = (I, (\actionset{\mu},
        \utility{\mu})_{\mu \in \mathcal{N}})$, where $I$ is the interaction graph,
        $\actionset{\mu}$ and $\utility{\mu}$ refer to $\mu$'s set of actions (with cardinality
        $|\actionset{\mu}| = n$) and utility function respectively.
  \end{definition}

  We consider the \emph{state} of agent $\mu$ to be the probability vector $\xmu \in
  \mathbb{R}^n$ where $\xmu_i$ is the probability with which the agent plays action $i$. It
  should be noted that this probability vector is often referred to as $\mu$'s \emph{mixed
  strategy}. With this in mind, we can construct, as their state space, the unit simplex on
  agent $\mu$'s action set. This is given as $\Delta_\mu := \{\xmu \in \mathbb{R}^n \, : \,
  \sum_i \xmuaction{i} = 1\}$. Also associated with each agent is a utility function

  The utility function, for each agent $\mu$, is given as $u^\mu(\xmu, \xnotmu)$ in which we use
  the standard notation $-\mu$ to refer to all agents other than $\mu$. Notice that this requires
  that each agent play the same strategy against all of their neighbours. What is unique about the
  NA game is the structure of the payoffs themselves. In this format, each agent is receives a
  reference $\sigma$ which is a weighted sum of each of their neighbours
  state. Formally
  %
  \begin{equation}
    \sigma^\mu = \sum_{\nu \in N^\mu} \wmunu \xnu.
  \end{equation}

  Then, the agent's must optimise their payoff with respect to this reference vector. Thus,
  instead of having to consider the actions of the entire population, or play individual games
  against each of their neighbours, the agent only has to consider $\sigma^\mu$ as a
  `measurement' of the local aggregate state and optimise with respect to this measurement.
  This allows us to make the reduction $u^\mu(\xmu, \xnotmu) = u^\mu(\xmu, \refmu)$. In
  particular, we consider that the agent is engaged in a matrix game against the
  reference vector so that
  %
  \begin{equation}
    u^\mu(\xmu, \refmu) = \xmu \cdot A^\mu \refmu = \xmu
                \cdot A^\mu \weightedsum.
  \end{equation}

  where $A^\mu$ is the payoff matrix associated to agent $\mu$. Note that this means we can write
  the game $\Gamma$ with the payoff matrices $A^mu$ in place of the utility functions
  $\utility{\mu}$. The NA game allows for the reduction of a multiplayer game into a series of
  two-player games. The agent's goal is to maximise their payoff with respect to the reference
  vector. As such, we define the best response correspondence $BR^\mu$ which maps any $\refmu$
  the set $\arg \max_{y \in \Delta_\mu} {u^\mu(y, \refmu)}$.

  Finally, a central concept of game theory is that of the Nash Equilibrium, in which no rational
  agent has the incentive to deviate from their current state. This can be formalised by saying
  that all agents are playing the best response to each other.

  This leads naturally to the definition of a Nash Equilibrium in an NA Game as

  \begin{definition}(NE) \label{def::NE}
    The set of vectors $\{ \NE{\mu}\}_{\mu \in \agentset}$ is an NE if, for all $\mu$,  
    \begin{equation*}
    \NE{\mu} \in BR^\mu (\refmu) = \arg \max_{x \in \Delta_\mu} u^\mu(x, w^{\mu \mu} x + \sum_{\nu \in N^\mu \backslash \{\bar{\mu}\}} \wmunu \NE{\nu}).
    \end{equation*} 
  \end{definition}

  \begin{remark}
    It can be seen that the concept of the Nash Equilibrium in the NA game is a natural
    extension of the NE in a bimatrix game $(A, B)$. In particular, if we consider an NA game
    with only two players and no self-loops then the above definition yields

    \begin{equation*}
      \NE{1} \in BR^1 (\sigma^1) = \arg \max_{x in \Delta_1} u^1 (x, \NE{2}), 
    \end{equation*}

    and similarly for $\NE{2}$. This is precisely the definition of an NE in a two player game \cite{}.
  \end{remark}

@@

  Finally we note that an NA game is \emph{zero-sum} if the utilities of each agent sum to zero for any strategy set $\{ \xmu \}_{\mu \in \agentset}$. Formally

  \begin{equation}
    \sum_\mu u^\mu(\xmu, \weightedsum) = 0.
  \end{equation}

>>>>

\fb{multi agent $\to$ multi-agent}

@@

NA games are a recent \cite{Parise2015} extension of
%introduction  which presents
%a variant of
aggregative games by adding an underlying structure to
the agent population. Since its introduction, distributed algorithms have
been built with the aim of finding NE in NA games. In particular,
\cite{Parise2015,Parise2020} consider the case in which payoffs are
given by Lipschitz functions with unique minimisers and apply standard
topological fixed-point towards designing algorithms that
converge to NE. We note that the \emph{Mann Iteration}, one of the
algorithms applied in \cite{Parise2020}, shares a remarkably similar
form with the discrete variant of FP. To the best of our knowledge,
this link has not been explored in the literature and is a venue for
research that we are currently pursuing. Another common algorithm for
distributed NE seeking is the projected gradient (resp. subgradient)
dynamics, which is explored in \cite{Zhang2020}
(resp.~\cite{Shokri2020, Shokri2021}). In all of these works, the cost
function is assumed to be convex, and therefore have a unique
minimiser. In fact this is a common assumption in works concerning NA
Games \cite{Zhu2021, Lei2020}, which we believe is due to its ubiquity
in control settings. We have not yet come across works which consider
NA games from the point of view of payoff matrices, which are more
common in multi-agent learning settings.  Furthermore, to the best of
our knowledge, this is the first work which introduces the application
of a strict \fb{?} learning algorithm in the NA setting.

\fb{the section on FP is very well-read, but maybe not everything is
  relevant for the following. I'd rather try to focus on what is
  relevant for our paper and contrast approaches.  Also, it is not
  clear what is the interest of FP for multi-agent learning.  }

\paragraph{Fictitious Play}
%
Fictitious Play \cite{BrownPublished,
  BrownUnpublished}
was introduced
%by Brown in 1951
as a 'natural' way
%in which
to approximate Nash Equilibris in zero-sum games. At the time, it was
shown that a discrete variant of FP
%the algorithm would
converge to 
NE in two-player zero-sum games \cite{Robinson}. Since then, a number
of results on convergence were proved in two-player games
\cite{Miyasawa, Polak, Berger, Monderer and Sela, Monderer and
  Shapley}, including
%as well as
corresponding convergence rates \cite{Harris,
  Shapiro}.
%Many results hold for both the discrete and continuous
%variants of FP.
The relation between the discrete and continuous variants of FP was
finally solved by Hofbauer \cite{Hofbauer}, who showed that attractors
of the discrete time process are also attractors for the continuous
time variant. As such, most of the analysis on FP focuses on the
continuous time variant \cite{Ostrovski}. It was later shown in
\cite{Shapley} that there is a family of games (now called the
\emph{Shapley family}) for which FP does not converge, but
rather shows periodic behaviour. \fb{we might remove the following: This led to research which looks at
this non convergent behaviour, specifically showing the existence of
more periodic orbits as well as chaotic behaviour
\cite{Sparrow}.} However, the work looking at fictitious play
%where there are
with more than two agents is sparse. In \cite{Sela}
%considers a game in which a
multi-player games are decomposed
%by requiring that each agent engage in a
in two-player games between each pair of players in the game.
%against each of the opponent.
Each agent's payoff is given by the sum of payoffs in all of these
subgames. It was found that, if this game is zero-sum, then FP will
converge. Similar results for more than two players were found for
games with identical interests \fb{?} in \cite{MondererShapley1994}. In
\cite{FPMFG}, the action of FP was considered in a mean field game,
and it was found again that the algorithm converges for zero-sum
games. The most general result, and the one most similar to our own,
appears in
%was found by Ewerhart and Valkanova in 2020
\cite{Ewerhart}, in which
the authors show that fictitious play converges in network games, where
each agent is engaged in a two-player game with each of their
neighbours. The rate of convergence was found to be independent of the
network size.  Our work extends the analysis of FP in multi-player
games by considering its action in NA Games, so that the agents do not
play individual games against each of their neighbours but rather a
single game against the aggregate of their neighbours. In particular
we also extend a result for two player games found by Ostrovski and
van Strein \cite{Payoff Performance} which showed that FP achieves
no-regret in the multi-player setting.

\section{Preliminaries}

\fb{I'd rather not leave an empty space between section and
  subsection.}

\subsection{Network Aggregative Games}
\label{sec::NAG}

The model we consider consists of
%is that there is
a set $\agentset = \{1,
           \ldots , N \}$ of agents, who are connected through an underlying
           interaction graph.
           %This graph is given by the tuple
          %$(\agentset, (\edgeset, \weightset))$ in which $\edgeset$ is
          %the set of all pairs $(\mu, \nu) \in \agentset \times \agentset$
           %connected.
           More formally:
%        
  \begin{definition}[Interaction Graph] \label{interactiongraph}
    Given a set $\agentset$ of agents, an {\em interaction graph} $I = (\agentset, (\edgeset,
    \weightset))$ is such that
    \begin{itemize}

    \item $\edgeset \subseteq \agentset^2$.  Then, the set of
      neighbours of agent $\mu$ is denoted as $N^\mu = \{\nu \in
      \agentset \mid (\mu, \nu) \in \edgeset\}$.

    \item $\weightset \in M_N(\mathbb{R})$ is the weighted \fb{or weight?} adjacency matrix, whose elements $w^{\mu
        \nu} \in [0, 1]$ expresses the importance that agent $\mu$ places on agent $\nu$. If $(\mu, \nu) \not
        \in \edgeset$ then $w^{\mu \nu} = 0$;        $\wmunu \in (0, 1]$ otherwise.
    \end{itemize}
  \end{definition}

  Given Def.~\ref{interactiongraph} of interaction graph,
  %With the above definition,
  we introduce network aggregative (NA) games:
%
  \begin{definition}[NA Game]
    A {\em network aggregative game} is
    %defined as the
a tuple $\Gamma = (I, (\actionset{\mu},
    \utility{\mu})_{\mu \in \mathcal{N}})$, where $I$ is an
    interaction graph, and for every agent $\mu \in \mathcal{N}$, $\actionset{\mu}$ and $\utility{\mu}$
    %refer to
    are $\mu$'s set of actions (with cardinality $|\actionset{\mu}| = n$)
    and utility function respectively.
  \end{definition}

  We define the \emph{state} of agent $\mu$ to be the probability
  vector $\xmu \in \mathbb{R}^n$, where $\xmu_i$ is the probability
  with which agent $\mu$ plays action $i$.
  %It should be noted that
  This
  probability vector is often referred to as $\mu$'s \emph{mixed
    strategy}. With this in mind, we can construct, as their state
  space, the {\em unit simplex} $\Delta_\mu$ on agent $\mu$'s action set, which
  is defined
  as $\Delta_\mu := \{\xmu \in \mathbb{R}^n \, : \, \sum_i
  \xmuaction{i} = 1\}$. Also associated with each agent is a utility
  function

  The utility, for each agent $\mu$ and action profile
  $(\xmu, \xnotmu)$, is given as $u^\mu(\xmu, \xnotmu)$ in which we
  use the standard notation $-\mu$ to refer to all agents other than
  $\mu$. Notice that this requires that each agent plays the same
  strategy against all of their neighbours. What is unique about 
  NA games is the structure of the payoffs themselves. In this format,
  each agent $\mu$ receives a reference $\sigma^{\mu} = \sum_{\nu \in N^\mu} \wmunu \xnu$, which is a weighted sum
  of each of their neighbours state.
  %Formally
  %
%  \begin{equation}
%    \sigma^\mu = \sum_{\nu \in N^\mu} \wmunu \xnu.
%  \end{equation}

  Then, the agent must optimise their payoff with respect to the
  reference vector $\sigma^{\mu}$. Thus, instead of considering
  the actions of the entire population, or play individual games
  against each of their neighbours, the agent only considers
  $\sigma^\mu$ as a `measurement' of the local aggregate state and
  optimises with respect to this measurement.  This allows us to make
  the reduction $u^\mu(\xmu, \xnotmu) = u^\mu(\xmu, \refmu)$. In
  particular, we consider that the agent is engaged in a matrix game
  against the reference vector so that
  %
  \begin{equation}
    u^\mu(\xmu, \refmu) = \xmu \cdot A^\mu \refmu = \xmu
                \cdot A^\mu \weightedsum.
  \end{equation}
%
  where $A^\mu$ is the payoff matrix associated with agent $\mu$. In particular,  this means we can write the game $\Gamma$ with the payoff
  matrices $A^\mu$ in place of the utility functions
  $\utility{\mu}$. Hence, the NA game allows for the reduction of a
  multi-player game into a series of two-player games. The agent's goal
  is to maximise their payoff with respect to the reference vector. As
  such, we define the best response correspondence $BR^\mu$, which maps
  every $\refmu$ to the set $\arg \max_{y \in \Delta_\mu} {u^\mu(y,
    \refmu)}$.

  Finally, a central concept of game theory is that of the Nash
  Equilibrium, in which no rational agent has the incentive to deviate
  from their current state. This can be formalised by saying that all
  agents are playing their best response to each other.
%
  This leads naturally to the definition of a Nash Equilibrium in an NA game as
%
  \begin{definition}(NE) \label{def::NE}
    The set of vectors $\{ \NE{\mu}\}_{\mu \in \agentset}$ is a {\em
      Nash equilibrium} if, for all $\mu$,
    \begin{equation*}
    \NE{\mu} \in BR^\mu (\refmu) = \arg \max_{x \in \Delta_\mu} u^\mu(x, w^{\mu \mu} x + \sum_{\nu \in N^\mu \backslash \{\bar{\mu}\}} \wmunu \NE{\nu}).
    \end{equation*} 
  \end{definition}

  \begin{remark}
    %It can be seen that
The notion of Nash equilibrium in the NA game is a natural
extension of the NE in a bimatrix game.
%$(A, B)$.
In particular, if we consider an NA game with only two players and no
self-loops then the above definition yields that $ \NE{1}$ is an NE iff
%
%    \begin{equation*}
 $     \NE{1} \in BR^1 (\sigma^1) = \arg \max_{x in \Delta_1} u^1 (x, \NE{2})$,
%    \end{equation*}
%
    and similarly for $\NE{2}$. This is precisely the definition of an NE in a two player game \cite{}.
  \end{remark}

@@

  Finally, we note that an NA game is \emph{zero-sum} if the utilities of each agent sum to zero for any strategy set $\{ \xmu \}_{\mu \in \agentset}$. Formally,
%
%  \begin{equation}
    $\sum_\mu u^\mu(\xmu, \weightedsum) = 0$.
%  \end{equation}
